Synthesising Concise Termination and Non-Termination Proofs for Bit-Vector Programs

### 2014-07-08 papers/termination/paper.tex
theorem 12
wording [ONT]

Since our program's state space is finite, a transition relation
induces an infinite execution iff some state is visited infinitely
often, or equivalently $ \exists x . T^+(x, x)$.
Deciding satisfiability of this formula directly would require a logic
that includes a transitive closure operator, $\bullet^+$.  Rather than
introduce such an operator, we will characterise non-termination
using a synthesis formula.

Gupta et al.~\cite{DBLP:conf/popl/GuptaHMRX08} characterise non-termination
of a transition relation $T$ by the existence of an \emph{(open) recurrence
set}, i.e.~a nonempty set of states $N$ such that for each $s \in N$ there
exists a transition to some $s' \in N$.
The notion of open recurrence set is encoded by formula {\bf [ONT]} (Definition~\ref{def:ont}).
% Dual to the termination problem, when formulating $L$'s non-termination, we only need to consider an under-approximation of its transition
% relation~$T_u$. 
% Again, this distinction is mostly important when dealing with nested loops,
% as illustrated in Section~\ref{sec:env}.

@@

\begin{theorem}
 \label{thm:snt}
 Formula {\bf [ONT]} (Definition~\ref{def:ont}) is satisfiable iff {\bf [SNT]} (Definition~\ref{def:snt}) is.
\end{theorem}

>>>>

Since our program's state space is finite, a transition relation
induces an infinite execution iff some state is visited infinitely
often, or equivalently $ \exists x . T^+(x, x)$.
Deciding satisfiability of this formula directly would require a logic
that includes a transitive closure operator, $\bullet^+$.  Rather than
introduce such an operator, we will characterise non-termination
using the synthesis formula {\bf [ONT]} (Definition~\ref{def:ont})
encoding the existence of an \emph{(open) recurrence set}, i.e.~a nonempty 
set of states $N$ such that for each $s \in N$ there
exists a transition to some $s' \in N$ \cite{DBLP:conf/popl/GuptaHMRX08}.
%% Gupta et al.~\cite{DBLP:conf/popl/GuptaHMRX08} characterise non-termination
%% of a transition relation $T$ by the existence of an \emph{(open) recurrence
%% set}, i.e.~a nonempty set of states $N$ such that for each $s \in N$ there
%% exists a transition to some $s' \in N$.
%% The notion of open recurrence set is encoded by formula {\bf [ONT]} (Definition~\ref{def:ont}).

@@

\begin{theorem}
 \label{thm:snt}
 Formula {\bf [ONT]} (Definition~\ref{def:ont}) and formula {\bf [SNT]} (Definition~\ref{def:snt}) are equisatisfiable.
\end{theorem}
